---
title: "Untitled"
format: html
---

This file contains the examples of functions and unit test by synthetic data

## Synthetic Data
We simulate synthetic yield curve data for two scenarios over two time horizons: 10 days and 960 days. The parameters for the Nelson-Siegel model are set as follows: beta_0 = 4, beta_1 = 0, beta_2 = -5, and lambda = 0.5, ensuring positive yields.

(Even though the 960-day case is exaggerated, it allows us to test the model's performance and correctness effectively.)

The two scenarios differ in the structure of the covariance matrix used for generating multivariate normal errors:
1. A diagonal covariance matrix (OLS case).
2. A non-diagonal covariance matrix (GLS case).

### Short time window OLS case
```{r}
synthetic_data_OLS_Short_R <- generate_data(T_ = 10, betas = c(5, -3, 3), lambda = 0.5, GLS = F)
synthetic_OLS_yields_Short_R <- synthetic_data_OLS_Short_R$yield_list

synthetic_data_OLS_Short_inverted <- generate_data(T_ = 10, betas = c(5, 0, 3), lambda = 0.5, GLS = F)
synthetic_OLS_yields_Short_inverted <- synthetic_data_OLS_Short_inverted$yield_list
```

### Long time window OLS case
```{r}
synthetic_data_OLS_Long_R <- generate_data(T_ = 960, betas = c(5, -3, 3), lambda = 0.5, GLS = F)
synthetic_OLS_yields_Long_R <- synthetic_data_OLS_Long_R$yield_list
```

### Short time window GLS case
```{r}
synthetic_data_GLS_Short_R <- generate_data(T_ = 10, betas = c(5, -3, 3), lambda = 0.5, GLS = T)
synthetic_GLS_yields_Short_R <- synthetic_data_GLS_Short_R$yield_list
```

### Long time window GLS case
```{r}
synthetic_data_GLS_Long_R <- generate_data(T_ = 960, betas = c(5, -3, 3), lambda = 0.5, GLS = T)
synthetic_GLS_yields_Long_R <- synthetic_data_GLS_Long_R$yield_list
```

```{r}
common_tenors <- c(1/12, 1/4, 1/2, 1,2,3,5,7,10,20)
```

## Data interpolation
Data interpolation for normal yield curve
```{r}
interp_NS_GLS <- interpolate_list(synthetic_GLS_yields_Long_R, 1, 10, 3)

plot(synthetic_GLS_yields_Long_R[[5]]$Maturity, 
  synthetic_GLS_yields_Long_R[[1]]$ZERO_YLD1, 
  xlab = "Time to Maturities", 
  ylab = "Yield Rate (Percent)", 
  main = "Interpolated Normal Yield Curve")
lines(interp_NS_GLS[[5]]$Maturity, interp_NS_GLS[[1]]$ZERO_YLD1)
```
Data interpolation for inverted yield curve
```{r}
interp_NS_OLS_inverted <- interpolate_list(synthetic_OLS_yields_Short_inverted, 1, 10, 3)

plot(synthetic_OLS_yields_Short_inverted[[5]]$Maturity, 
  synthetic_OLS_yields_Short_inverted[[1]]$ZERO_YLD1, 
  xlab = "Time to Maturities", 
  ylab = "Yield Rate (Percent)", 
  main = "Interpolated InvertedYield Curve")
lines(interp_NS_OLS_inverted[[5]]$Maturity, interp_NS_OLS_inverted[[1]]$ZERO_YLD1)
```

## Estimation

### fit_nelson_siegel in OLS case over short time window
```{r}
OLS_fit_Short_R <- fit_nelson_siegel(
  synthetic_OLS_yields_Short_R, 
  lambda = 0.5, 
  start = 1, 
  tenors = common_tenors, 
  T_ = 10
)

cat("Short-term OLS Estimated Betas:", OLS_fit_Short_R$betas, "\n")
cat("Short-term OLS Estimated Variance:", OLS_fit_Short_R$sigma2, "\n")
cat("Short-term True Variance:", synthetic_data_OLS_Short_R$sigma2, "\n")
```
The short-term OLS fit demonstrates accurate parameter estimation, with betas close to the true values and a variance estimate closely aligning with the synthetic variance.

### fit_nelson_siegel in OLS case over long time window
```{r}
OLS_fit_Long_R <- fit_nelson_siegel(
  synthetic_OLS_yields_Long_R,
  lambda = 0.5, 
  start = 1, 
  tenors = common_tenors, 
  T_ = 960
)

cat("Long-term OLS Estimated Betas:", OLS_fit_Long_R$betas, "\n")
cat("Long-term OLS Estimated Variance:", OLS_fit_Long_R$sigma2, "\n")
cat("Long-term True Variance:", synthetic_data_OLS_Long_R$sigma2, "\n")

```
The long-term OLS fit achieves highly accurate parameter estimation, with betas nearly matching the true values and the variance estimate closely approximating the synthetic variance.

### fit_nelson_siegel in GLS case over short time window
```{r}
GLS_fit_Short_R<- fit_nelson_siegel_GLS(
  synthetic_GLS_yields_Short_R, 
  lambda = 0.5, 
  start = 1, 
  tenors = common_tenors, 
  T_ = 10
)

cat("Short-term GLS Estimated Betas:", GLS_fit_Short_R$betas, "\n")
GLS_fit_Short_R$cov_mat
synthetic_data_GLS_Short_R$cov_mat

```
Fr short time window, the beta estimate is decent, while the estimated covaraince matrix does not perform well.


### fit_nelson_siegel in GLS case over long time window

```{r}

GLS_fit_Long_R <- fit_nelson_siegel_GLS(
  synthetic_GLS_yields_Long_R, 
  lambda = 0.5, 
  start = 1, 
  tenors = common_tenors, 
  T_ = 960
)

cat("Long-term GLS Estimated Betas:", GLS_fit_Long_R$betas, "\n")
GLS_fit_Long_R$cov_mat
synthetic_data_GLS_Long_R$cov_mat
```
When the time window gets larger, the estimated covariance matrix converges to the true covariance matrix.

### get_likelihood in OLS case (estimation for lambda) 

```{r}
get_likelihood(yield_list = synthetic_GLS_yields_Short_R, 
               lambda_list = seq(0.1, .8, 0.01), 
               start = 1, 
               T_ = 10, 
               GLS = T, 
               tenors = common_tenors)
```
The true lambda equals to 0.5, indicating that the estimation is not really accurate in short time window.

```{r}
get_likelihood(yield_list = synthetic_GLS_yields_Long_R, 
               lambda_list = seq(0.1, 1, 0.01), 
               start = 1, 
               T_ = 960, 
               GLS = T, 
               tenors = common_tenors)
```
With longer time period, the estimation of lambda is accurate enough

### fit_NS_paramters 

```{r}
fit_NS_parameters(yield_list = synthetic_GLS_yields_Long_R, 
                  lambda_list = seq(0.1, 1, 0.01), 
                  GLS = T,
                  start = 1, 
                  T_ = 960, 
                  tenors = common_tenors)
```
The function returns best lambda and betas.