---
title: "Testing goodness of fit for different time window lengths"
author: "Manny"
date: "2024-11-27"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will generate synthetic data and estimate the NS parameters over a time window to compare goodness-of-fit.

The first will be done by generating yield curves over a time window with constant true parameters.

We will first estimate the parameters using OLS over the entire time window. 
Then, we will estimate the parameters again but using OLS every year.
Then, we will do the same but every month.
Finally, we will do every week.

We will use AIC to compare goodness of fit across 3 simulated scenarios & one real scenario

```{r}
cs_tenors_r <- c(1/2, 5, 7, 10, 20)
cs_tenors_i <- c(1/4, 5, 10, 15, 20)
# These betas we will use for simulations case 2 and 3
periodic_betas <- matrix(
  c(5, 5, 4, 5, 5, 5, 5, 4, 5, 5, 5, 5, 4, 5, 5, 5, 5,  # L_t
    -3, -3, 0, 2, 4, -3, -3, 0, 2, 4, -3, -3, 0, 2, 4, -3, -3,  # S_t
    3, 0, -3, 2, -2, 3, 0, -3, 2, -2, 3, 0, -3, 2, -2, 3, 0  # C_t
  ),
  nrow = 3,  # 4 rows (including headers for clarity)
  byrow = TRUE
)

```

Generate yield curves


```{r cars}
set.seed(123)
Case1_Data <- generate_data(T_ = 960, betas = c(5,-3,3), lambda= 0.35, GLS = F)$yield_list
Case2_Data = list()
for(i in 1:16){
  add_df <- generate_data(T_ = 60, betas = periodic_betas[,i], lambda = 0.35,
                GLS = F, seed = 123)$yield_list
  Case2_Data <- append(Case2_Data, add_df)
}

Case3_Data <- last960FredYields
```



Interpolate yields (this will take a few minutes so only run once)

```{r}
interp_Case1 <- interpolate_list(Case1_Data, 1, 960)
interp_Case2 <- interpolate_list(Case2_Data, 1, 960)
interp_Case3 <- interpolate_list(Case3_Data, 1, 960) # this takes about 10-15 minutes to run
```

Take a look and see if the bootstrap seems reasonable (no crazy jumps and such.)
```{r}
plot(Case3_Data[[200]])
lines(interp_Case3[[200]])
```

Over each year (assume 250 trading days a year):
This function compure the AIC values for each fit
```{r}
compute_aic <- function(yield_list, window_length, tenors, GLS = F){
  # yield_list: Parameter of the form of a list of data frames containing ZCB spot rate
  # window_length: length of time window
  # tenors: list of time to maturities
  # GLS: whether use GLS algorithm
  
  total_T <- length(yield_list)
  T_ <- window_length 
  N <- length(tenors) 
  nw <- ceiling(round(total_T / T_, 3)) # Number of windows
  betas_matrix <- matrix(nrow = 3, ncol = nw)
  aic_vec <- c()
  lambda_vec <- c()
  if(nw > 1){
    for(i in 1:(nw - 1)){
      NS_fit <- fit_NS_parameters(yield_list = yield_list,
                                        lambda_list = seq(0.05, 0.95, 0.025),
                                        GLS = GLS,
                                        start = (i - 1) * T_ + 1,
                                        T_ = T_,
                                        tenors = tenors) # we fit nelson Siegel paramters
      betas_matrix[,i] <- NS_fit$betas
      aic_vec[i] <- 2 * 3 - 2 * NS_fit$log_likelihood # compute the AIC value
      lambda_vec[i] <- NS_fit$lambda
    } 
    if(total_T - (nw-1)*T_ == T_){ # check if the remaining data points perfectly match the window_length (whether the last window is a complete window)
      NS_fit <- fit_NS_parameters(yield_list = yield_list,
                                          lambda_list = seq(0.05, 0.95, 0.025),
                                          GLS = GLS,
                                          start = (nw-1) * T_ + 1,
                                          T_ = T_,
                                          tenors = tenors)
      betas_matrix[,nw] <- NS_fit$betas
      aic_vec[nw] <- 2 * 3 - 2 * NS_fit$log_likelihood
      lambda_vec[nw] <- NS_fit$lambda
    } else { # processes an incomplete last window. (avoid subscript out of bound)
      print((nw-1) * T_ + 1)
      NS_fit <- fit_NS_parameters(yield_list = yield_list,
                                          lambda_list = seq(0.05, 0.95, 0.025),
                                          GLS = GLS,
                                          start = (nw-1) * T_ + 1,
                                          T_ = total_T - (nw-1) * T_,
                                          tenors = tenors)
      betas_matrix[,nw] <- NS_fit$betas
      aic_vec[nw] <- 2 * 3 - 2 * NS_fit$log_likelihood
      lambda_vec[nw] <- NS_fit$lambda
    }
  }
  else{
    NS_fit <- fit_NS_parameters(yield_list = yield_list,
                                          lambda_list = seq(0.05, 0.95, 0.025),
                                          GLS = GLS,
                                          start = 1,
                                          T_ = T_,
                                          tenors = tenors)
    betas_matrix[,1] <- NS_fit$betas
    aic_vec[1] <- 2 * 3 - 2 * NS_fit$log_likelihood
    lambda_vec[1] <- NS_fit$lambda
  }
  return(list(aic = aic_vec, betas = betas_matrix, lambda = lambda_vec))
}

```

## OLS
Case 1, OLS
```{r}
set.seed(123)
aic_1_5 <- compute_aic(Case1_Data, 5, cs_tenors_r, GLS = T)
aic_1_10 <- compute_aic(Case1_Data, 10, cs_tenors_r, GLS = T)
aic_1_20 <- compute_aic(Case1_Data, 20, cs_tenors_r, GLS = T)
aic_1_40 <- compute_aic(Case1_Data, 40, cs_tenors_r, GLS = T)
aic_1_60 <- compute_aic(Case1_Data, 60, cs_tenors_r, GLS =T)
aic_1_80 <- compute_aic(Case1_Data, 80, cs_tenors_r, GLS =T)
aic_1_120 <- compute_aic(Case1_Data, 120, cs_tenors_r, GLS = T)


aic_df1 <- data.frame(Time_Window = c(rep(5, ceiling(960/5)),
                                   rep(10, ceiling(960/10)),
                                   rep(20, ceiling(960/20)),
                                   rep(40, ceiling(960/40)),
                                   rep(60, ceiling(960/60)),
                                   rep(80, ceiling(960/80)),
                                   rep(120, ceiling(960/120))),
          
                                 AIC_per_window = c(aic_1_5$aic,
                                                    aic_1_10$aic,
                                                    aic_1_20$aic,
                                                    aic_1_40$aic,
                                                    aic_1_60$aic,
                                                    aic_1_80$aic,
                                                    aic_1_120$aic
                                                    ),
                     lambda_per_window = c(aic_1_5$lambda,
                                                    aic_1_10$lambda,
                                                    aic_1_20$lambda,
                                                    aic_1_40$lambda,
                                                    aic_1_60$lambda,
                                                    aic_1_80$lambda,
                                                    aic_1_120$lambda
                                                    ))


aic_df1$Time_Window <- as.factor(aic_df1$Time_Window)

bpAic1 <- ggplot(aic_df1, aes(x = Time_Window, y = AIC_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "AIC") + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

bpLogLik1 <- ggplot(aic_df1, aes(x = Time_Window, y = lambda_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "Lambda MLE") + geom_hline(yintercept = 0.35, linetype = 'longdash', color = 'red') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))
makeLatexLabs(bpAic1, newY = '$AIC(\\hat{\\beta})$', newTitle = "AIC of Estimates per Time Window Length")
makeLatexLabs(bpLogLik1, newY = '$\\hat{\\lambda$}^{MLE}', newTitle = "MLE of $\\lambda$ per Time Window Length")
```
For some date, lets see what the nelson siegel fitted line looks like compared to the real yields to maturity

```{r}
NS_C1 <- fit_NS_parameters(Case1_Data, GLS = F, start = 1, T_ = 120, tenors = cs_tenors_r)
betasC1 <- NS_C1$betas
lamC1 <- NS_C1$lambda

gridMonthTo20 <- seq(1/12, 20, 0.1)
phi2c1 <- (1 - exp(- lamC1 * gridMonthTo20)) / (lamC1 * gridMonthTo20)
phi3c1 <- phi2c1 - exp(- lamC1 * gridMonthTo20)
trueLine <- data.frame(x = gridMonthTo20,
                       y = betasC1[1] + betasC1[2] * phi2c1 + betasC1[3] * phi3c1,
                       source = 'Fitted Line')

trueLinePlot <- ggplot() + 
  geom_point(data = Case1_Data[[1]], aes(x = Maturity, y = ZERO_YLD1)) + 
  geom_line(data = trueLine, aes(x = x, y = y, color = source), linetype = 'longdash') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

makeLatexLabs(trueLinePlot, '$\\tau$  (years)', '$y_t(\\tau)$  (Annualized  %)', "Simulated Data + Nelson Siegel Fitted Line (Case 1)")
```
Case 2, OLS
```{r}
aic_2_5 <- compute_aic(Case2_Data, 5, cs_tenors_r, GLS = F)
aic_2_10 <- compute_aic(Case2_Data, 10, cs_tenors_r, GLS = F)
aic_2_20 <- compute_aic(Case2_Data, 20, cs_tenors_r, GLS = F)
aic_2_40 <- compute_aic(Case2_Data, 40, cs_tenors_r, GLS = F)
aic_2_60 <- compute_aic(Case2_Data, 60, cs_tenors_r, GLS = F)
aic_2_80 <- compute_aic(Case2_Data, 80, cs_tenors_r, GLS = F)
aic_2_120 <- compute_aic(Case2_Data, 120, cs_tenors_r, GLS = F)

aic_df2 <- data.frame(Time_Window = c(
    rep(5, ceiling(960/5)),
    rep(10, ceiling(960/10)),
    rep(20, ceiling(960/20)),
    rep(40, ceiling(960/40)),
    rep(60, ceiling(960/60)),
    rep(80, ceiling(960/80)),
    rep(120, ceiling(960/120))),
  AIC_per_window = c(
    aic_2_5$aic,
    aic_2_10$aic,
    aic_2_20$aic,
    aic_2_40$aic,
    aic_2_60$aic,
    aic_2_80$aic,
    aic_2_120$aic), 
  lambda_per_window = c(
    aic_2_5$lambda,
    aic_2_10$lambda,
    aic_2_20$lambda,
    aic_2_40$lambda,
    aic_2_60$lambda,
    aic_2_80$lambda,
    aic_2_120$lambda))

aic_df2$Time_Window <- as.factor(aic_df2$Time_Window)

bpAic2 <- ggplot(aic_df2, aes(x = Time_Window, y = AIC_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "AIC") + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

bpLam2 <- ggplot(aic_df2, aes(x = Time_Window, y = lambda_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "Lambda MLE") + geom_hline(yintercept = 0.35, linetype = 'longdash', color = 'red') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))
makeLatexLabs(bpAic2, newY = '$AIC(\\hat{\\beta})$', newTitle = "AIC of Estimates per Time Window Length")
makeLatexLabs(bpLam2, newY = '$\\hat{\\lambda$}^{MLE}', newTitle = "MLE of $\\lambda$ per Time Window Length")


```
```{r}
NS_C2 <- fit_NS_parameters(Case2_Data, GLS = F, start = 41, T_ = 40, tenors = cs_tenors_r)
betasC2 <- NS_C2$betas
lamC2 <- NS_C2$lambda

gridMonthTo20 <- seq(1/12, 20, 0.1)
phi2c2 <- (1 - exp(- lamC2 * gridMonthTo20)) / (lamC2 * gridMonthTo20)
phi3c2 <- phi2c2 - exp(- lamC2 * gridMonthTo20)
trueLine <- data.frame(x = gridMonthTo20,
                       y = betasC2[1] + betasC2[2] * phi2c2 + betasC2[3] * phi3c2,
                       source = 'Fitted Line')

trueLinePlot <- ggplot() + 
  geom_point(data = Case2_Data[[59]], aes(x = Maturity, y = ZERO_YLD1)) + 
  geom_line(data = trueLine, aes(x = x, y = y, color = source), linetype = 'longdash') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

makeLatexLabs(trueLinePlot, '$\\tau$  (years)', '$y_t(\\tau)$  (Annualized  %)', "Simulated Data + Nelson Siegel Fitted Line (Case 1)")
```
Case 3, OLS
```{r}
#aic_3_5 <- compute_aic(Case3_Data, 5, cs_tenors_r, GLS = T)
aic_3_10 <- compute_aic(Case3_Data, 10, cs_tenors_r, GLS = F)
aic_3_20 <- compute_aic(Case3_Data, 20, cs_tenors_r, GLS = F)
aic_3_40 <- compute_aic(Case3_Data, 40, cs_tenors_r, GLS = F)
aic_3_60 <- compute_aic(Case3_Data, 60, cs_tenors_r, GLS = F)
aic_3_80 <- compute_aic(Case3_Data, 80, cs_tenors_r, GLS = F)
aic_3_120 <- compute_aic(Case3_Data, 120, cs_tenors_r, GLS = F)

aic_df3 <- data.frame(Time_Window = c(#rep(5, ceiling(960/5)),
                                   rep(10, ceiling(960/10)),
                                   rep(20, ceiling(960/20)),
                                   rep(40, ceiling(960/40)),
                                   rep(60, ceiling(960/60)),
                                   rep(80, ceiling(960/80)),
                                   rep(120, ceiling(960/120))),
          
                                 AIC_per_window = c(#aic_3_5$aic,
                                                    aic_3_10$aic,
                                                    aic_3_20$aic,
                                                    aic_3_40$aic,
                                                    aic_3_60$aic,
                                                    aic_3_80$aic,
                                                    aic_3_120$aic
                                                    ),
                     lambda_per_window = c(#aic_3_5$lambda,
                                                    aic_3_10$lambda,
                                                    aic_3_20$lambda,
                                                    aic_3_40$lambda,
                                                    aic_3_60$lambda,
                                                    aic_3_80$lambda,
                                                    aic_3_120$lambda
                                                    ))


aic_df3$Time_Window <- as.factor(aic_df3$Time_Window)

bpAic3 <- ggplot(aic_df3, aes(x = Time_Window, y = AIC_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "AIC") + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

bpLam3 <- ggplot(aic_df3, aes(x = Time_Window, y = lambda_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "Lambda MLE") + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

makeLatexLabs(bpAic3, newY = '$AIC(\\hat{\\beta})$', newTitle = "AIC of Estimates per Time Window Length")
makeLatexLabs(bpLam3, newY = '$\\hat{\\lambda$}^{MLE}', newTitle = "MLE of $\\lambda$ per Time Window Length")


```


```{r}

NS_C3 <- fit_NS_parameters(Case3_Data, GLS = F, start = 800, T_ = 80, tenors = cs_tenors_r)
betasC3 <- NS_C3$betas
lamC3 <- NS_C3$lambda

gridMonthTo20 <- seq(1/12, 20, 0.1)
phi2c1 <- (1 - exp(- lamC3 * gridMonthTo20)) / (lamC3 * gridMonthTo20)
phi3c1 <- phi2c1 - exp(- lamC3 * gridMonthTo20)
trueLine <- data.frame(x = gridMonthTo20,
                       y = betasC3[1] + betasC3[2] * phi2c1 + betasC3[3] * phi3c1,
                       source = 'Fitted Line')

trueLinePlot <- ggplot() + 
  geom_point(data = Case3_Data[[890]][1:10,], aes(x = Maturity, y = ZERO_YLD1)) + 
  geom_line(data = trueLine, aes(x = x, y = y, color = source), linetype = 'longdash') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

makeLatexLabs(trueLinePlot, '$\\tau$  (years)', '$y_t(\\tau)$  (Annualized  %)', "Simulated Data + Nelson Siegel Fitted Line (Case 1)")
```


## GLS

Case 1, GLS
```{r}
aic_1_10 <- compute_aic(Case1_Data, 10, cs_tenors_r, GLS = T)
aic_1_20 <- compute_aic(Case1_Data, 20, cs_tenors_r, GLS = T)
aic_1_40 <- compute_aic(Case1_Data, 40, cs_tenors_r, GLS = T)
aic_1_60 <- compute_aic(Case1_Data, 60, cs_tenors_r, GLS =T)
aic_1_80 <- compute_aic(Case1_Data, 80, cs_tenors_r, GLS =T)
aic_1_120 <- compute_aic(Case1_Data, 120, cs_tenors_r, GLS = T)


aic_df1 <- data.frame(Time_Window = c(
                                   rep(10, ceiling(960/10)),
                                   rep(20, ceiling(960/20)),
                                   rep(40, ceiling(960/40)),
                                   rep(60, ceiling(960/60)),
                                   rep(80, ceiling(960/80)),
                                   rep(120, ceiling(960/120))),
          
                                 AIC_per_window = c(
                                                    aic_1_10$aic,
                                                    aic_1_20$aic,
                                                    aic_1_40$aic,
                                                    aic_1_60$aic,
                                                    aic_1_80$aic,
                                                    aic_1_120$aic
                                                    ),
                     lambda_per_window = c(
                                                    aic_1_10$lambda,
                                                    aic_1_20$lambda,
                                                    aic_1_40$lambda,
                                                    aic_1_60$lambda,
                                                    aic_1_80$lambda,
                                                    aic_1_120$lambda
                                                    ))


aic_df1$Time_Window <- as.factor(aic_df1$Time_Window)

bpAic1 <- ggplot(aic_df1, aes(x = Time_Window, y = AIC_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "AIC") + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

bpLogLik1 <- ggplot(aic_df1, aes(x = Time_Window, y = lambda_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "Lambda MLE") + geom_hline(yintercept = 0.35, linetype = 'longdash', color = 'red') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))
makeLatexLabs(bpAic1, newY = '$AIC(\\hat{\\beta})$', newTitle = "AIC of Estimates per Time Window Length")
makeLatexLabs(bpLogLik1, newY = '$\\hat{\\lambda$}^{MLE}', newTitle = "MLE of $\\lambda$ per Time Window Length")
```

For some date, lets see what the nelson siegel fitted line looks like compared to the real yields to maturity

```{r}
NS_C1 <- fit_NS_parameters(Case1_Data, GLS = F, start = 1, T_ = 120, tenors = cs_tenors_r)
betasC1 <- NS_C1$betas
lamC1 <- NS_C1$lambda

gridMonthTo20 <- seq(1/12, 20, 0.1)
phi2c1 <- (1 - exp(- lamC1 * gridMonthTo20)) / (lamC1 * gridMonthTo20)
phi3c1 <- phi2c1 - exp(- lamC1 * gridMonthTo20)
trueLine <- data.frame(x = gridMonthTo20,
                       y = betasC1[1] + betasC1[2] * phi2c1 + betasC1[3] * phi3c1,
                       source = 'Fitted Line')

trueLinePlot <- ggplot() + 
  geom_point(data = Case1_Data[[1]], aes(x = Maturity, y = ZERO_YLD1)) + 
  geom_line(data = trueLine, aes(x = x, y = y, color = source), linetype = 'longdash') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

makeLatexLabs(trueLinePlot, '$\\tau$  (years)', '$y_t(\\tau)$  (Annualized  %)', "Simulated Data + Nelson Siegel Fitted Line (Case 1)")
```

Case 2, GLS

```{r}
aic_2_10 <- compute_aic(Case2_Data, 10, cs_tenors_r, GLS = T)
aic_2_20 <- compute_aic(Case2_Data, 20, cs_tenors_r, GLS = T)
aic_2_40 <- compute_aic(Case2_Data, 40, cs_tenors_r, GLS = T)
aic_2_60 <- compute_aic(Case2_Data, 60, cs_tenors_r, GLS = T)
aic_2_80 <- compute_aic(Case2_Data, 80, cs_tenors_r, GLS = T)
aic_2_120 <- compute_aic(Case2_Data, 120, cs_tenors_r, GLS = T)

aic_df2 <- data.frame(Time_Window = c(
    rep(10, ceiling(960/10)),
    rep(20, ceiling(960/20)),
    rep(40, ceiling(960/40)),
    rep(60, ceiling(960/60)),
    rep(80, ceiling(960/80)),
    rep(120, ceiling(960/120))),
  AIC_per_window = c(
    aic_2_10$aic,
    aic_2_20$aic,
    aic_2_40$aic,
    aic_2_60$aic,
    aic_2_80$aic,
    aic_2_120$aic), 
  lambda_per_window = c(
    aic_2_10$lambda,
    aic_2_20$lambda,
    aic_2_40$lambda,
    aic_2_60$lambda,
    aic_2_80$lambda,
    aic_2_120$lambda))

aic_df2$Time_Window <- as.factor(aic_df2$Time_Window)

bpAic2 <- ggplot(aic_df2, aes(x = Time_Window, y = AIC_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "AIC") + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

bpLam2 <- ggplot(aic_df2, aes(x = Time_Window, y = lambda_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "Lambda MLE") + geom_hline(yintercept = 0.35, linetype = 'longdash', color = 'red') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))
makeLatexLabs(bpAic2, newY = '$AIC(\\hat{\\beta})$', newTitle = "AIC of Estimates per Time Window Length")
makeLatexLabs(bpLam2, newY = '$\\hat{\\lambda$}^{MLE}', newTitle = "MLE of $\\lambda$ per Time Window Length")


```


```{r}
NS_C2 <- fit_NS_parameters(Case2_Data, GLS = F, start = 41, T_ = 40, tenors = cs_tenors_r)
betasC2 <- NS_C2$betas
lamC2 <- NS_C2$lambda

gridMonthTo20 <- seq(1/12, 20, 0.1)
phi2c2 <- (1 - exp(- lamC2 * gridMonthTo20)) / (lamC2 * gridMonthTo20)
phi3c2 <- phi2c2 - exp(- lamC2 * gridMonthTo20)
trueLine <- data.frame(x = gridMonthTo20,
                       y = betasC2[1] + betasC2[2] * phi2c2 + betasC2[3] * phi3c2,
                       source = 'Fitted Line')

trueLinePlot <- ggplot() + 
  geom_point(data = Case2_Data[[59]], aes(x = Maturity, y = ZERO_YLD1)) + 
  geom_line(data = trueLine, aes(x = x, y = y, color = source), linetype = 'longdash') + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

makeLatexLabs(trueLinePlot, '$\\tau$  (years)', '$y_t(\\tau)$  (Annualized  %)', "Simulated Data + Nelson Siegel Fitted Line (Case 1)")
```

Case 3, GLS

```{r}
aic_3_10 <- compute_aic(Case3_Data, 10, cs_tenors_r, GLS = T)
aic_3_20 <- compute_aic(Case3_Data, 20, cs_tenors_r, GLS = T)
aic_3_40 <- compute_aic(Case3_Data, 40, cs_tenors_r, GLS = T)
aic_3_60 <- compute_aic(Case3_Data, 60, cs_tenors_r, GLS = T)
aic_3_80 <- compute_aic(Case3_Data, 80, cs_tenors_r, GLS = T)
aic_3_120 <- compute_aic(Case3_Data, 120, cs_tenors_r, GLS = T)

aic_df3 <- data.frame(Time_Window = c(
                                   rep(10, ceiling(960/10)),
                                   rep(20, ceiling(960/20)),
                                   rep(40, ceiling(960/40)),
                                   rep(60, ceiling(960/60)),
                                   rep(80, ceiling(960/80)),
                                   rep(120, ceiling(960/120))),
          
                                 AIC_per_window = c(
                                                    aic_3_10$aic,
                                                    aic_3_20$aic,
                                                    aic_3_40$aic,
                                                    aic_3_60$aic,
                                                    aic_3_80$aic,
                                                    aic_3_120$aic
                                                    ),
                     lambda_per_window = c(
                                                    aic_3_10$lambda,
                                                    aic_3_20$lambda,
                                                    aic_3_40$lambda,
                                                    aic_3_60$lambda,
                                                    aic_3_80$lambda,
                                                    aic_3_120$lambda
                                                    ))


aic_df3$Time_Window <- as.factor(aic_df3$Time_Window)

bpAic3 <- ggplot(aic_df3, aes(x = Time_Window, y = AIC_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "AIC") + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

bpLam3 <- ggplot(aic_df3, aes(x = Time_Window, y = lambda_per_window)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(x = "Time Window Length (Days)",
       y = "Lambda MLE") + theme_economist_white() + theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5, margin = margin(b= 10)))

makeLatexLabs(bpAic3, newY = '$AIC(\\hat{\\beta})$', newTitle = "AIC of Estimates per Time Window Length")
makeLatexLabs(bpLam3, newY = '$\\hat{\\lambda$}^{MLE}', newTitle = "MLE of $\\lambda$ per Time Window Length")


```
